Rank Verification for Exponential Families

### 2016-07-08 verifying-winner.tex
minor fixes

\begin{proof}
Under $H_0$, one of the $H_{0j}$ is true. Suppose this is $H_{01}$. In this case,
$$\PP\left(p_* \le \alpha\right) \le \PP\left(p_1 \le \alpha\right) \le \alpha.$$
And thus $p_*$ is a valid $p$-value for the union null hypothesis.

\end{proof}

@@

\begin{proof}
The type I error --- the probability of incorrectly declaring anyone other than the best to be the best --- is
\[
\PP\left[\bigcup_{j \text{ not best}} \text{declare } j \text{ best}\right] \leq \sum_j \PP\left[ \text{declare } j \text{ best} \mid j \text{ wins}\right] \PP[j \text{ wins}].
\]
Because only one candidate can win (recall that ties are broken randomly), it is sufficient to show for each $j$ that 
$\PP_{H_{0j}}\left[ \text{reject } H_{0j} \mid j \text{ wins}\right] \leq \alpha$.

@@

Finally, by construction, $p_{1j}$ satisfies
$$\PP_{H_{01j}}\left(p_{1j} < \alpha \;\middle|\; M_{1j}, X_{\setminus\left\{1, j\right\}}, A_1\right) = \alpha.$$
Marginalizing over $M_{1j}, X_{\setminus\left\{1, j\right\}}$,
$$\PP_{H_{01j}}\left(p_{1j} < \alpha \;\middle|\; A_1\right) = \alpha.$$
Therefore these $p_{1j}$ are indeed valid $p$-values.

@@

Now suppose the exponential family is defined on the counting measure instead. If ties are broken independently and randomly, the end points on the rays can be considered as ``half an atom'' if the coordinates are integers, and a ``full atom'' otherwise.
\WFcomment{It also relies on the atoms on the two rays ``matching up,'' right?}
\end{proof}

@@

The selective type I error we opted to control for Procedure 1 offers the conventional, marginal type I error rate.
\begin{equation}
\PP_{H_{01}}\left(\text{reject } H_{01}\right) = \PP_{H_{01}}\left(\text{Winner is } 1\right) \PP_{H_{01}}\left(\text{reject } H_{01} \;\middle|\; \text{Winner is } 1\right) \le \frac{n-1}{n} \alpha,
\label{eqn:marginal}
\end{equation}
because, under the hypothesis $H_{01}$, the probability of $1$, which is not the best population, winning is at most $\frac{n-1}{n}$. Selective type I error rate control provides more than just marginal type I error rate control. \citet{Fithian:2015uj} has shown that it helps controlling familywise error rate, which we will utilize in \Cref{sec:stepdown}.

@@

\begin{proof}
It suffices to prove the ``if'' part, as the ``only if'' part can be follows from swapping the role of $\theta_1$ and $\theta_2$. For any fixed $a$, and $X_1 \ge a$ and $X_2 < a$, we have $X_1 > X_2$ and
$$\exp\left(\theta_1 X_1 + \theta_2 X_2 + \cdots + \theta_n X_n - \psi\left(\theta\right)\right) g\left(X\right) \ge \exp\left(\theta_1 X_2 + \theta_2 X_1 + \cdots + \theta_n X_n - \psi\left(\theta\right)\right) g\left(X\right).$$
Integrating both sides over the region $\left\{X: X_1 \ge a, X_2 < a\right\}$ gives
$$\PP\left(X_1 \ge a, X_2 < a\right) \ge \PP\left(X_1 < a, X_2 \ge a\right).$$
Now adding $\PP\left(X_1 \ge a, X_2 \ge a\right)$ to both probabilities gives
$$\PP\left(X_1 \ge a\right) \ge \PP\left(X_2 \ge a\right),$$
meaning that $X_1$ is greater than $X_2$ in stochastic order.
\end{proof}

@@

\citet{Gupta:1967wg} devises a rule to select a subset that includes the maximum $\pi_j$. In other words, if the selected subset is $J\left(X\right)$, it guarantees
\begin{equation}\label{eq:subsetsel}
\PP\left(\argmax_j \pi_j \in J\left(X\right)\right) \ge 1 - \alpha.
\end{equation}
This is achieved by finding an integer $d$, as a function on $m$, $n$ and $\alpha$, and selecting the subset
$$J\left(X\right) = \left\{j: X_j \ge \max_k X_k - d\right\}.$$
Here, we take $d(m, n, \alpha)$ to be the smallest integer such that \eqref{eq:subsetsel} holds for any $\pi$.

@@

A fundamental difference between the two methods is the definition of the threshold. The threshold from \citet{Gupta:1967wg} depends solely on the setup of the model: $m$ and $n$. Our threshold, however, depends on the data itself. Such difference is amplified when $n$ is large. When $n$ is large, $d$ from \citeauthor{Gupta:1967wg} is computed by considering the often-worst case scenario where $\pi = \left(\frac{1}{2}, \frac{1}{2}, 0, \ldots, 0\right)$. Hence $d$ is the $1-\alpha$ \WFcomment{$1-\alpha/2$?} quantile of
$$X_1 - X_2 \sim m - 2 \cdot \text{Binomial}\left(m, \frac{1}{2}\right) \approx \text{Normal}\left(0, m\right)$$
and $d \approx \sqrt{m} z_\alpha$ \WFcomment{where $z_\alpha$ is BLAH}. On the other hand, our method defines a threshold based on the $1 - \frac{n}{n-1} \cdot \frac{\alpha}{2}$ quantile of
$$X_1 - X_2 \mid X_1 + X_2 \sim X_1 + X_2 - 2 \cdot \text{Binomial}\left(X_1 + X_2, \frac{1}{2}\right),$$
which is approximately $\sqrt{X_1 + X_2} z_{\alpha / 2}$. The smaller $\left(X_1 + X_2\right) / m$ is (e.g.\ if $\pi_1 + \pi_2$ is smaller), the more different the two thresholds are, suggesting that our test is more powerful.

@@

Formally, we wish to find the biggest $j$ such that we can infer
$$\theta_{[1]} > \ldots > \theta_{[j]} > \max_{k>j} \theta_{[k]}.$$
If $j_0$ is the true value of $j$, the FWER refers to $\PP\left(j \ne j_0\right)$.

@@

Hence similar to the proof of Part 1 of \Cref{thm:main}, Schur-concavity ensures $p_{j\left(j+1\right)} \ge p_{jk}$ for all $k>j$, meaning that it is sufficient to compare $X_j$ and $X_{j+1}$. In fact, we can take $p_j = p_{j\left(j+1\right)}$. These are valid selective $p$-value as required in \citet{Fithian:2015uj}, as
$$\PP_\theta \left(p_j \le \alpha \mid \theta_{[k]} \text{ is not the } k\text{-th largest, for some } k \le j\right) \le \alpha \text{ a.s.}, ~~~~~~~~ \text{for all } \theta,$$
by tower property. Note that the conditioning is indeed a random event as the order of $\theta$ depends on $X$ (emphasized with the square brackets here). \KHcomment{I should probably add a line transitioning from using $\mathcal{M}$ to suddenly mentioning $\theta$ here, shouldn't I?}

@@

The obvious error in the authors' reasoning is perhaps understandable since there is, presently, no agreed-upon statistical solution for this commonplace problem. Using our approach, they could have performed a two-tailed binomial test of the winner (``Fame'') vs. the runner-up (``Benevolence''). This test returns a $p$-value of
$$p = 2\PP\left[\text{Binomial}\left(8 + 5, 0.5\right) \ge 8\right] = 0.58,$$
indicating that their data did not support their major finding.

>>>>

\begin{proof}
Under $H_0$, one of the $H_{0j}$ is true. Suppose this is $H_{01}$. In this case,
$$\PP\left[p_* \le \alpha\right] \le \PP\left[p_1 \le \alpha\right] \le \alpha.$$
And thus $p_*$ is a valid $p$-value for the union null hypothesis.
\end{proof}

@@

\begin{proof}
The type I error --- the probability of incorrectly declaring any non-best to be the best --- is
\begin{equation}
\PP\left[\bigcup_{j \ne j^*} \text{declare } j \text{ best}\right] \le \sum_{j \ne j^*} \PP\left[\text{declare } j \text{ best} \;\middle|\; j \text{ wins}\right] \PP\left[j \text{ wins}\right],
\label{eq:marginal}
\end{equation}
where $j^*$ is the best (fixed when the distribution is determined). The probability above is taken over all distributions where $j^*$ is the best. However, since there is exactly one winner at all times (recall that ties are broken at all times), it is sufficient to bound $\PP_{j^* \text{ is the best}}\left[ \text{declare } j \text{ best} \;\middle|\; j \text{ wins}\right]$, for each $j \ne j^*$. Since $j^*$ being the best implies that $j$ is not the best, it is, in turn, sufficient to show that
$$\PP_{j \text{ is not the best}}\left[ \text{declare } j \text{ best} \;\middle|\; j \text{ wins}\right].$$

@@

Finally, by construction, $p_{1j}$ satisfies
$$\PP_{H_{01j}}\left[p_{1j} < \alpha \;\middle|\; M_{1j}, X_{\setminus\left\{1, j\right\}}, A_1\right] = \alpha.$$
Marginalizing over $M_{1j}, X_{\setminus\left\{1, j\right\}}$,
$$\PP_{H_{01j}}\left[p_{1j} < \alpha \;\middle|\; A_1\right] = \alpha.$$
Therefore these $p_{1j}$ are indeed valid $p$-values.

@@

Now suppose the exponential family is defined on the counting measure instead. If ties are broken independently and randomly, the end points on the rays can be considered as ``half an atom'' if the coordinates are integers, and a ``full atom'' otherwise. The number of atoms on each ray is the same (after the extension $\tilde{A}$) and thus the atoms on each ray can be paired up. The inequalities above still hold for each pair of the atoms.
\end{proof}

@@

The selective type I error we opted to control for Procedure 1 offers the conventional, marginal type I error rate, as shown in \eqref{eq:marginal}. More precisely with our notation in the proof,
\begin{equation}
\PP_{H_{01}}\left[\text{reject } H_{01}\right] = \PP_{H_{01}}\left[\text{Winner is } 1\right] \PP_{H_{01}}\left[\text{reject } H_{01} \;\middle|\; \text{Winner is } 1\right] \le \frac{n-1}{n} \alpha,
\label{eqn:marginal}
\end{equation}
because, under the hypothesis $H_{01}$, the probability of $1$, which is not the best population, winning is at most $\frac{n-1}{n}$. Selective type I error rate control provides more than just marginal type I error rate control. \citet{Fithian:2015uj} has shown that it helps controlling familywise error rate, which we will utilize in \Cref{sec:stepdown}.

@@

\begin{proof}
It suffices to prove the ``if'' part, as the ``only if'' part can be follows from swapping the role of $\theta_1$ and $\theta_2$. For any fixed $a$, and $X_1 \ge a$ and $X_2 < a$, we have $X_1 > X_2$ and
$$\exp\left(\theta_1 X_1 + \theta_2 X_2 + \cdots + \theta_n X_n - \psi\left(\theta\right)\right) g\left(X\right) \ge \exp\left(\theta_1 X_2 + \theta_2 X_1 + \cdots + \theta_n X_n - \psi\left(\theta\right)\right) g\left(X\right).$$
Integrating both sides over the region $\left\{X: X_1 \ge a, X_2 < a\right\}$ gives
$$\PP\left[X_1 \ge a, X_2 < a\right] \ge \PP\left[X_1 < a, X_2 \ge a\right].$$
Now adding $\PP\left[X_1 \ge a, X_2 \ge a\right]$ to both probabilities gives
$$\PP\left[X_1 \ge a\right] \ge \PP\left[X_2 \ge a\right],$$
meaning that $X_1$ is greater than $X_2$ in stochastic order.
\end{proof}

@@

\citet{Gupta:1967wg} devises a rule to select a subset that includes the maximum $\pi_j$. In other words, if the selected subset is $J\left(X\right)$, it guarantees
\begin{equation}\label{eq:subsetsel}
\PP\left[\argmax_j \pi_j \in J\left(X\right)\right] \ge 1 - \alpha.
\end{equation}
This is achieved by finding an integer $d$, as a function on $m$, $n$ and $\alpha$, and selecting the subset
$$J\left(X\right) = \left\{j: X_j \ge \max_k X_k - d\right\}.$$
Here, we take $d(m, n, \alpha)$ to be the smallest integer such that \eqref{eq:subsetsel} holds for any $\pi$.

@@

A fundamental difference between the two methods is the definition of the threshold. The threshold from \citet{Gupta:1967wg} depends solely on the setup of the model: $m$ and $n$. Our threshold, however, depends on the data itself. Such difference is amplified when $n$ is large. When $n$ is large, $d$ from \citeauthor{Gupta:1967wg} is computed by considering the often-worst case scenario where $\pi = \left(\frac{1}{2}, \frac{1}{2}, 0, \ldots, 0\right)$. Hence $d$ is the upper $\alpha$ quantile of
$$X_1 - X_2 \sim m - 2 \cdot \text{Binomial}\left(m, \frac{1}{2}\right) \approx \text{Normal}\left(0, m\right)$$
and $d \approx \sqrt{m} z_\alpha$, where $z_\alpha$ is the upper $\alpha$ quantile of a standard Gaussian. On the other hand, our method defines a threshold based on the upper $\frac{n}{n-1} \cdot \frac{\alpha}{2}$ quantile of
$$X_1 - X_2 \mid X_1 + X_2 \sim X_1 + X_2 - 2 \cdot \text{Binomial}\left(X_1 + X_2, \frac{1}{2}\right),$$
which is approximately $\sqrt{X_1 + X_2} z_{\alpha / 2}$. The smaller $\left(X_1 + X_2\right) / m$ is (e.g.\ if $\pi_1 + \pi_2$ is smaller), the more different the two thresholds are, suggesting that our test is more powerful.

@@

Formally, we wish to find the biggest $j$ such that we can infer
$$\theta_{[1]} > \ldots > \theta_{[j]} > \max_{k>j} \theta_{[k]}.$$
If $j_0$ is the true value of $j$, the FWER refers to $\PP\left[j \ne j_0\right]$.

@@

Hence similar to the proof of Part 1 of \Cref{thm:main}, Schur-concavity ensures $p_{j\left(j+1\right)} \ge p_{jk}$ for all $k>j$, meaning that it is sufficient to compare $X_j$ and $X_{j+1}$. In fact, we can take $p_j = p_{j\left(j+1\right)}$. These are valid selective $p$-value as required in \citet{Fithian:2015uj}, as
$$\PP_\theta \left[p_j \le \alpha \mid \theta_{[k]} \text{ is not the } k\text{-th largest, for some } k \le j\right] \le \alpha \text{ a.s.}, ~~~~~~~~ \text{for all } \theta,$$
by tower property. Note that the conditioning is indeed a random event as the order of $\theta$ depends on $X$ (emphasized with the square brackets here). \KHcomment{I should probably add a line transitioning from using $\mathcal{M}$ to suddenly mentioning $\theta$ here, shouldn't I?}

@@

The obvious error in the authors' reasoning is perhaps understandable since there is, presently, no agreed-upon statistical solution for this commonplace problem. Using our approach, they could have performed a two-tailed binomial test of the winner (``Fame'') vs. the runner-up (``Benevolence''). This test returns a $p$-value of
$$p = 2 \cdot \PP\left[\text{Binomial}\left(8 + 5, 0.5\right) \ge 8\right] = 0.58,$$
indicating that their data did not support their major finding.
